---
title: "Assignment 4"
author: "Júlia Papa"
date: "2025-07-20"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# dataset upload
library(stargazer)
library(readr)
qog <- read_csv("qog_long.csv")
```

**Instructions:** Use the previous qog_long.csv dataset on E-learning to do the following tasks

**Exercise 1:** Dichotomize the governance quality score (iiag_gov), using "50" as the cutoff (coded 1 when the variable is higher than 50, and 0 otherwise)

```{r ex1, include=TRUE}
qog$high_gov_quality <- as.integer(qog$iiag_gov > 50)
# Test to verify if it worked
head(qog$high_gov_quality)
```

**Exercise 2:** Run a logit model on this binary outcome, with three explanatory variables, including mad_gdppc (GDP per capita), al_ethnic (ethnic fragmentation), and bti_ffe (free and fair election score)

```{r ex2, include=TRUE}
logit_model <- glm(high_gov_quality ~ mad_gdppc + al_ethnic + bti_ffe,
                   data = qog,
                   family = binomial(link = "logit"))
logit_model
```

**Exercise 3:** Report Regression result with a table

```{r ex3, include = TRUE}
library(stargazer)
stargazer(logit_model, type = "text")
```

**Exercise 4:** Interpret the magnitude of effect for statistically significant variable(s), using both "divide-by-4'' and odds methods.

bti_ffe is the only significant variable (p \< 0.01)\*\*\*

```{r ex4, include = TRUE}
# Extract coefficient for bti_ffe
coeff_bti_ffe <- coef(logit_model)["bti_ffe"]

# Interpretation using "divide-by-4" rule
effect_bti_ffe_divide_by_4 <- coeff_bti_ffe / 4
effect_bti_ffe_divide_by_4

# Interpretation using Odds Ratios
odds_ratio_bti_ffe <- exp(coeff_bti_ffe)
odds_ratio_bti_ffe
```

-   **Divide-by-4 rule interpretation**

    -   For every one-unit increase in bti_ffe, the probability of high governance quality is estimated to increase by approximately 0.1707087;

-   **Odds ratios interpretation**

    -    For every one-unit increase in bti_ffe, the odds of high governance quality are multiplied by approximately 1.979481 (an increase of about 97.94809).
