---
title: "Assignment 2"
author: "Júlia Papa"
date: "2025-07-08"
output: word_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stargazer)
library(readr)
qog <- read_csv("C:/Users/julia/Documents/FGV/Advanced_Methods/Aula 2/qog_long.csv")
View(qog)
```

Use the following variables from qog_long.csv to conduct regression
analysis and answer the following questions (answers, table, and figures
are put in one word document):

-   iiag_gov: overall quality of governance (outcome variable) •

-   mad_gdppc: GDP per capita

-   pwt_pop: population in millions

-   al_language: language fragmentation

-   bti_ffe: free and fair election, 10 for the best

-   al_ethnic: ethnic fractionalization

**EXERCISE 1** - Built a linear model with all the variables, estimate
parameters, create a table, and interpret the result for each
explanatory variable (4 pts)

```{r ex1, include = TRUE}
model1 <- lm(iiag_gov ~ mad_gdppc + pwt_pop + al_language + al_ethnic + bti_ffe, data = qog)
model1
## I'll use stargazer to create the table
 stargazer(model1, type = "text", title = "Regression Results for Governance Quality",
         dep.var.labels = "Overall Quality of Governance",
           covariate.labels = c("GDP per capita", "Population (millions)",
                                "Language Fragmentation", "Ethnic Fractionalization",
                                "Free and Fair Election"),
           notes = "Standard errors in parentheses. * p < 0.1; ** p < 0.05; *** p < 0.01")
```

**RESULTS INTERPRETATION FOR EACH EXPLANATORY VARIABLE**

-   **mad_gdppc**: A one-unit increase in GDP per capita is associated
    with a 0.001 unit increase in governance quality, a small, but
    statistically significant positive effect.

-   **pwt_pop**: Population size shows a minor negative association
    (-0.029) with governance quality, but this effect is not
    statistically significant.

-   **al_language**: Language fragmentation has a positive coefficient
    (6.735) but is not statistically significant, suggesting no clear
    linear impact on governance quality.

-   **bti_ffe**: A one-unit increase in the free and fair election
    indicator significantly increases governance quality by 2.589 units,
    indicating a strong positive relationship.

-   **al_ethnic**: Ethnic fractionalization has a negative association
    (-5.563) with governance quality, but this is not statistically
    significant.

-   **Constant**: The estimated baseline governance quality score is
    32.557 when all other variables are zero, and this is statistically
    significant.

**EXERCISE 2 -** According to the model, how would the quality of
governance score change if the free and fair election indicator
increases from 4 to 6? (2 pt)

It would change by 5.177504 points.

```{r ex2, include= TRUE}
beta_ffe <- coef(model1)["bti_ffe"]
change_in_ffe <- 6 - 4
delta <- beta_ffe * change_in_ffe
print(delta)
```

**EXERCISE 3 -** Among the explanatory variables with significance,
which one has a bigger impact? (2 pt)

bti_ffe is the one tha has a bigger impact.

**EXERCISE 4 -** Do you think the model works fine with the data
according to the model diagnostics? Briefly explain your reasons (2 pt)

```{r ex4, include=TRUE, echo=TRUE}
summary(model1)
plot(model1)
```

The model **does not work perfectly fine** with the data. While it shows
**strong explanatory power** (R²=0.717, significant F-statistic), the
diagnostic plots indicate issues. There is **evidence of
heteroscedasticity** in the Residuals vs. Fitted and Scale-Location
plots, where the spread of residuals increases with fitted values. The
Normal Q-Q plot also shows **slight deviations from normality** at the
tails. These violations of assumptions mean the standard errors and
p-values might not be entirely reliable, even though no highly
influential observations were found.
