---
title: "Assignment 1 - Júlia Papa"
author: "Júlia Papa"
date: "2025-07-08"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
qog_2019 <- read_csv("C:/Users/julia/Documents/FGV/Advanced_Methods/Aula 1/qog_2019.csv")
View(qog_2019)
```

**Exercise 1** - 1. Briefly explain the relationships between the critical value, p-value, and statistical significance (4 pts)

These 3 concepts are fundamental to hypothesis testing:

### 1. P-value and Statistical Significance ($\alpha$ )

-   When the p-value is \< or = to the $\alpha$, the result is considered statistically significant, meaning there is strong evidence to reject H0.

-   When the p-value is \> than $\alpha$, the result is not statistically significat, meaning there is not enough evidence to reject H0.

### 2. Critical Value and Statistical Significance ($\alpha$)

-   The critical value is directly determined by the chosen significance level For a given statistical test and degrees of freedom, a specific $\alpha$ will correspond to a particular critical value(s) that marks the rejection region(s). A smaller $\alpha$ (e.g., 0.01 instead of 0.05) will lead to more extreme critical values, making it harder to reject H0, thus requiring stronger evidence.

### 3. P-value and Critical Value

-   These two approaches lead to the same conclusion regarding the H0:

    -   When you use the critical value approach, you calculate a test statistic and you calculate a test statistic from your data and compare it to the critical value. If the test statistic falls in the rejection region, you reject the H0.

    -   When you use the p-value approach, you calculate the p-value associated with your test statistic. If this p-value is less than $\alpha$ it implies that your test statistic is in the rejection region defined by the critical value.

**Exercise 2 -** How many countries that have a "g_effective" score higher than China? (2 pt)

52 countries have a g_effective score higher than China.

```{r ex2, include=TRUE}
china <- qog_2019$g_effective[qog_2019$cname == "China"]
higher_china <- sum(qog_2019$g_effective > china, na.rm = T)
higher_china
```

**Exercise 3 -** What are the mean and standard deviation of "g_effective" score of these countries found in the last question? (2 pt)

The mean is 1.256738 and the standard deviation is 0.4827013.

```{r ex3, include = TRUE}
higher <- qog_2019[qog_2019$g_effective > china,] 
# mean 
mean_ge <- mean(higher$g_effective, na.rm = T)
mean_ge
# standard deviation
sd_ge <- sd(higher$g_effective, na.rm = T)
sd_ge
```

**Exercise 4 -** If China's polity2 score increases by 8 units, what would be the mean of GDP per capita for the countries whose polity2 score is higher than China's? (2 pt)

The mean of GDP per capita for these countries would be 16532.66

```{r ex4, include = TRUE}

qog_2019$polity2_new <- qog_2019$polity2
qog_2019$polity2_new[qog_2019$cname == 'China'] <-  qog_2019$polity2[qog_2019$cname == 'China']
china_polity2_new <- qog_2019$polity2_new[qog_2019$cname == "China"]
higher_polity2 <- qog_2019[qog_2019$polity2_new > china_polity2_new, ]
mean_polity2 <- mean(higher_polity2$gdppc, na.rm = T)
mean_polity2
```
